Borel regularity for ODEs and thimble integrals

### 2023-05-01 exp-resurgence.tex
Fix typo in norm bound for algebraic integral operators
Veronica, thanks for spotting the typo independently. Now I'm sure
it wasn't just something I was too tired to understand.
updates

\title{Borel regularity for ODEs and exponential integrals}
\author{Veronica Fantini and Aaron Fenyes}

@@

The asymptotics of $f$ at the starting point of $\Gamma_{\zeta, \alpha}$ control the asymptotics of $\laplace_{\zeta, \alpha} f$ at the infinite end of $\Gamma_{z, 0}$. Once we see how this works for $\alpha = 0$, Section~\ref{translation} will do the rest. Let $F = \laplace_{\zeta, 0} f$. Equation~1.8 of \ref{laplace-tfm} shows\footnote{The argument cited still works in our generality. For holomorphic $f$, one can also use Equation 1.5 of {\em Borel-Laplace Transform and Asymptotic Theory} \textcolor{magenta}{(Sternin \& Shatalov)}.} that
\[ f \in O_{\zeta, 0 \leftarrow}(1) \quad\Longrightarrow\quad F \in O_{z, 0 \rightarrow}\left(\frac{1}{z}\right). \]
More generally, for $\tau > -1$ \textcolor{magenta}{[prove or cite]},
\[ f \in O_{\zeta, 0 \leftarrow}(\zeta^\tau) \quad\Longrightarrow\quad F \in O_{z, 0 \rightarrow}\left(\frac{1}{z^{1 + \tau}}\right). \]
Exact power law asymptotics relate similarly \textcolor{magenta}{[prove or cite]}:
\[ f \sim \zeta^\tau \text{ at the start of } \Gamma_{\zeta, 0} \quad\Longrightarrow\quad F \sim \frac{\Gamma(1+\tau)}{z^{1+\tau}} \text{ at the end of } \Gamma_{z, 0}. \]
\textcolor{magenta}{[The big-$O$ asymptotics dictionary is interesting, but we might not need it. Consider dropping.]}
\subsubsection{Action on integral operators}\label{L-int-op}
When $\varphi \in \mathcal{E}_\zeta$, you can use differentiation under the integral to show that~\cite[Theorem~1.34]{laplace-tfm}
\begin{equation}%%\label{id:L-mult}
\laplace_\zeta (\zeta^n \varphi) = \big({-\tfrac{\partial}{\partial z}}\big)^n \laplace_\zeta \varphi.
\end{equation}
for all integers $n \ge 0$. You can also use a 2d integration argument, akin to the one in \cite[Theorem~2.39]{laplace-tfm}, to show that $\fracderiv{-\lambda}{\zeta}{a} \varphi \in \mathcal{E}_\zeta$ and
\[ \laplace_{\zeta, a}\,\fracderiv{-\lambda}{\zeta}{a} \varphi = z^{-\lambda} \laplace_{\zeta, a} \varphi \]
for all $\lambda \in (0, \infty)$.
\subsubsection{Change of translation chart}\label{translation}
Define a new coordinate $\zeta_\alpha$ on $\C$ so that $\zeta = \alpha + \zeta_\alpha$. From the calculation
\begin{align*}
\laplace_{\zeta, a} \varphi & = \int_{\Gamma_{\zeta, \alpha}} e^{-z \zeta}\,\varphi\;d\zeta \\
& = \int_{\Gamma_{\zeta_\alpha, 0}} e^{-z(\alpha + \zeta_\alpha)}\,\varphi\;d\zeta_\alpha \\
& = e^{-\alpha z} \int_{\Gamma_{\zeta_\alpha, 0}} e^{-z\zeta_\alpha}\,\varphi\;d\zeta_\alpha \\
& = e^{-\alpha z} \laplace_{\zeta_\alpha, 0} \varphi,
\end{align*}
we learn that
\[ \laplace_{\zeta_\alpha, 0} \varphi = e^{\alpha z} \laplace_{\zeta, \alpha} \varphi. \]
\subsubsection{Rescaling of translation structure}
Let's rescale the translation structure of $\C$, expanding displacements by a factor of $\mu \in (0, \infty)$. The coordinate $\xi = \mu\zeta$ is a chart for the new translation structure. The corresponding frequency coordinate $x \maps T^*\C \to \C$ is given by $d\xi \mapsto 1$, so $x = \mu^{-1} z$. From the calculation
\begin{align*}
\laplace_{\xi, 0} \varphi & = \int_{\Gamma_{\xi, 0}} e^{-x\xi}\,\varphi\;d\xi \\
& = \int_{\Gamma_{\zeta, 0}} e^{-z \zeta}\,\varphi\;\mu\,d\zeta \\
& = \mu\,\laplace_{\zeta, 0} \varphi
\end{align*}
we learn that
\[ \laplace_{\xi, 0} \varphi = \mu\,\laplace_{\zeta, 0} \varphi. \]
Note that $\laplace_{\xi, 0}$ is defined in the new translation structure on $\C$, while $\laplace_{\zeta, 0}$ is defined in the old translation structure. We can still compare them, because they both turn complex-valued functions on $\C$ into holomorphic functions on $T^*\C$.

@@

\begin{itemize}
\item $P$ is a monic degree-$d$ polynomial with simple zeros at $-\alpha_j$ for $j=1,...,d$ and such that $P'(-\alpha_j)\neq 0$ for every $j=1,...,d$;
\item $Q$ is a degree-$(d-1)$ polynomial and $Q(-\alpha_j)\neq 0$ for all $j=1,...,d$;
\item $R_0, \ldots, R_d$ are entire functions of exponential type
\end{itemize}

@@

\begin{equation}
\hat{P}_j:=P(-\zeta_j)+\partial_{\zeta_j}^{-1}\circ Q(-\zeta_j)+\partial_{\zeta_j}^{-2}\circ\left[R_0(\partial_{\zeta_j}^{-1})+R_1(\partial_{\zeta_j}^{-1})\circ (-\zeta_j)+...+R_d(\partial_{\zeta_j}^{-1})\circ (-\zeta_j)^d\right]
\end{equation}

such that $\psi_j$ blows-up as $\zeta^{\tau_j-1}$ at $\alpha_j$:
${\psi}_j(\zeta_j)=\zeta_j^{-\tau_j-1}+f_j$ and $f_j\in\mathcal{HL}^{\infty,-\tau_j}(\Omega)$. 
\end{theorem}
\begin{proof}
It is enough to show that $\hat{P}_j$ satisfies the assumptions of Theorem 2.1.3 of {\tt regular singular Volterra equations}.  In fact, $P(-\zeta_j)$ is a polynomial, also defined at $\zeta_j=0$ and by assumption $P'(0)\neq 0$. $Q$ is also a polynomial defined at $\zeta_j=0$. Finally, assuming $R_k(z)=\sum_{j\geq 0} R_{k,j}z^j$

@@

\begin{theorem}\label{thm2-dim}
For every $j=1,...,d$ let $\psi_j(\zeta_j)$ be the unique solution of $\hat{P}_j\psi_j=0$ as in in Theorem \ref{thm1-dim}. Then $\psi_j(\zeta_j)$ admits a well defined Laplace transform $\Psi_j:=\mathcal{L}_{\zeta_j}^{\theta_j}$ for some direction $\theta_j$.
Furthermore, $\tilde{\Psi}_j$ is Borel summable and its Borel sum is proportional to $\Psi_j$. 
\end{theorem}
\begin{proof}
We divide the proof in the following steps:
\begin{itemize}
\item[$(a)$] prove that $\psi_j(\zeta_j)$ admits Laplace transform 
\item[$(b)$] let $\hat{\psi}_j$ be the sum of the germ $\tilde{\psi}_j:=\mathcal{B}\Psi_j$; prove that $\hat{\psi}_j \propto \psi_j$ 
\item[$(c)$] prove that $\mathcal{L}_{\zeta_j}^{\theta_j}\circ\mathcal{B}\series{\Psi}_j\propto \Psi_j $. 
\end{itemize}
\end{proof}

>>>>

\title{Borel regularity for ODEs and thimble integrals}
\author{Veronica Fantini and Aaron Fenyes}

@@

The asymptotics of $f$ at the starting point of $\Gamma_{\zeta, \alpha}$ control the asymptotics of $\laplace_{\zeta, \alpha} f$ at the infinite end of $\Gamma_{z, 0}$. Once we see how this works for $\alpha = 0$, Section~\ref{translation} will do the rest. Let $F = \laplace_{\zeta, 0} f$. Equation~1.8 of \ref{laplace-tfm} shows\footnote{The argument cited still works in our generality. For holomorphic $f$, one can also use Equation 1.5 of {\em Borel-Laplace Transform and Asymptotic Theory} \textcolor{magenta}{(Sternin \& Shatalov)}.} that
\[ f \in O_{\zeta, 0 \leftarrow}(1) \quad\Longrightarrow\quad F \in O_{z, 0 \rightarrow}\left(\frac{1}{z}\right). \]
More generally, for $\tau > -1$ \textcolor{magenta}{[prove or cite]},
\[ f \in O_{\zeta, 0 \leftarrow}(\zeta^\tau) \quad\Longrightarrow\quad F \in O_{z, 0 \rightarrow}\left(\frac{1}{z^{1 + \tau}}\right). \]
Exact power law asymptotics relate similarly \textcolor{magenta}{[prove or cite]}:
\[ f \sim \zeta^\tau \text{ at the start of } \Gamma_{\zeta, 0} \quad\Longrightarrow\quad F \sim \frac{\Gamma(1+\tau)}{z^{1+\tau}} \text{ at the end of } \Gamma_{z, 0}. \]
\textcolor{magenta}{[The big-$O$ asymptotics dictionary is interesting, but we might not need it. Consider dropping.]}
\subsubsection{Action on integral operators}\label{L-int-op}
When $\varphi \in \mathcal{E}_\zeta$, you can use differentiation under the integral to show that~\cite[Theorem~1.34]{laplace-tfm}
\begin{equation}%%\label{id:L-mult}
\laplace_\zeta (\zeta^n \varphi) = \big({-\tfrac{\partial}{\partial z}}\big)^n \laplace_\zeta \varphi.
\end{equation}
for all integers $n \ge 0$. You can also use a 2d integration argument, akin to the one in \cite[Theorem~2.39]{laplace-tfm}, to show that $\fracderiv{-\lambda}{\zeta}{a} \varphi \in \mathcal{E}_\zeta$ and
\[ \laplace_{\zeta, a}\,\fracderiv{-\lambda}{\zeta}{a} \varphi = z^{-\lambda} \laplace_{\zeta, a} \varphi \]
for all $\lambda \in (0, \infty)$.
\subsubsection{Change of translation chart}\label{translation}
Define a new coordinate $\zeta_\alpha$ on $\C$ so that $\zeta = \alpha + \zeta_\alpha$. From the calculation
\begin{align*}
\laplace_{\zeta, a} \varphi & = \int_{\Gamma_{\zeta, \alpha}} e^{-z \zeta}\,\varphi\;d\zeta \\
& = \int_{\Gamma_{\zeta_\alpha, 0}} e^{-z(\alpha + \zeta_\alpha)}\,\varphi\;d\zeta_\alpha \\
& = e^{-\alpha z} \int_{\Gamma_{\zeta_\alpha, 0}} e^{-z\zeta_\alpha}\,\varphi\;d\zeta_\alpha \\
& = e^{-\alpha z} \laplace_{\zeta_\alpha, 0} \varphi,
\end{align*}
we learn that
\[ \laplace_{\zeta_\alpha, 0} \varphi = e^{\alpha z} \laplace_{\zeta, \alpha} \varphi. \]

\color{magenta}

Write explicit definition $\mathcal{L}_{\zeta_{\alpha},0}^{\theta}\varphi$ specifying the argument of $\varphi$ in the integration.

\begin{equation}
\mathcal{L}_{\zeta_\alpha,0}^{\theta}\varphi:=\int_{\Gamma_{\zeta_\alpha,0}^{\theta}}e^{-z\zeta_\alpha}\varphi(\zeta_\alpha-\alpha)d\zeta_\alpha
\end{equation}

\subsubsection{Rescaling of translation structure}
Let's rescale the translation structure of $\C$, expanding displacements by a factor of $\mu \in (0, \infty)$. The coordinate $\xi = \mu\zeta$ is a chart for the new translation structure. The corresponding frequency coordinate $x \maps T^*\C \to \C$ is given by $d\xi \mapsto 1$, so $x = \mu^{-1} z$. From the calculation
\begin{align*}
\laplace_{\xi, 0} \varphi & = \int_{\Gamma_{\xi, 0}} e^{-x\xi}\,\varphi\;d\xi \\
& = \int_{\Gamma_{\zeta, 0}} e^{-z \zeta}\,\varphi\;\mu\,d\zeta \\
& = \mu\,\laplace_{\zeta, 0} \varphi
\end{align*}
we learn that
\[ \laplace_{\xi, 0} \varphi = \mu\,\laplace_{\zeta, 0} \varphi. \]
Note that $\laplace_{\xi, 0}$ is defined in the new translation structure on $\C$, while $\laplace_{\zeta, 0}$ is defined in the old translation structure. We can still compare them, because they both turn complex-valued functions on $\C$ into holomorphic functions on $T^*\C$.

@@

\begin{itemize}
\item $P$ is a monic degree-$d$ polynomial with simple zeros at $\alpha_j$ for $j=1,...,d$ and such that $P'(-\alpha_j)\neq 0$ for every $j=1,...,d$;
\item $Q$ is a degree-$(d-1)$ polynomial and $Q(-\alpha_j)\neq 0$ for all $j=1,...,d$;
\item $R_0, \ldots, R_d$ are entire functions of exponential type
\end{itemize}

@@

\color{brown}
\begin{equation}
\hat{P}_j:=P(-\zeta_j)+\partial_{\zeta_j}^{-1}\circ Q(-\zeta_j)+\partial_{\zeta_j}^{-2}\circ\left[R_0(\partial_{\zeta_j}^{-1})+R_1(\partial_{\zeta_j}^{-1})\circ (-\zeta_j)+...+R_d(\partial_{\zeta_j}^{-1})\circ (-\zeta_j)^d\right]
\end{equation}
\color{black}
such that $\psi_j$ blows-up as $\zeta^{\tau_j-1}$ at $\alpha_j$:
${\psi}_j(\zeta_j)=\zeta_j^{-\tau_j-1}+f_j$ and $f_j\in\mathcal{HL}^{\infty,-\tau_j}(\Omega)$. 
\end{theorem}
\begin{proof}
It is enough to show that $\hat{P}_j$ satisfies the assumptions of Theorem 2.1.3 of {\tt regular singular Volterra equations}.  In fact, $P(-\zeta_j)$ is a polynomial, also defined at $\zeta_j=0$ and by assumption $P'(0)\neq 0$. $Q$ is also a polynomial defined at $\zeta_j=0$. Finally, assuming $R_k(z)=\sum_{j\geq 0} R_{k,j}z^j$

@@

\begin{theorem}\label{thm2-dim}
For every $j=1,...,d$ let $\psi_j(\zeta_j)$ be the unique solution of $\hat{P}_j\psi_j=0$ as in in Theorem \ref{thm1-dim}. Then $\psi_j(\zeta_j)$ admits a well defined Laplace transform $\Psi_j:=\mathcal{L}_{\zeta_j}^{\theta_j}$ for some direction $\theta_j$.
Furthermore, let $\tilde{\Psi}_j(z)=e^{-\alpha_j z}z^{-\tau_j}\sum_{n=0}^{\infty}a_{j,n}z^{-n}$ be the formal solution of $\mathcal{P}\Psi=0$ with $a_{j,0}=1$ for every $j=1,...,d$. Then $\series{\Psi}_j(z)$ is Borel summable and its Borel sum is proportional to $\Psi_j$. 
\end{theorem}
\begin{proof}
We divide the proof in the following steps:
\begin{itemize}
\item[$(a)$] prove that $\psi_j(\zeta_j)$ admits Laplace transform 
\item[$(b)$] let $\hat{\psi}_j$ be the sum of the germ $\series{\psi}_j:=\mathcal{B}\series{\Psi}_j$; prove that $\hat{\psi}_j \propto \psi_j$ 
\item[$(c)$] prove that $\mathcal{L}_{\zeta_j}^{\theta_j}\circ\mathcal{B}\series{\Psi}_j\propto \Psi_j $. 
\end{itemize}

Step $(c)$ clearly follows from steps $(a)$ and $(b)$. We now prove step $(b)$: first notice that if $\series{\Psi}_j$ solves $\mathcal{P}\Psi=0$ then $\sum_{n=0}^{\infty}a_{j,n}z^{-n-\tau_j}$ is a solution of $\mathcal{P}_j\Psi=0$ with

\begin{equation}
\mathcal{P}_j:=P(-\alpha_j+\partial_z)+\frac{1}{z}Q(-\alpha_j+\partial_z)+\frac{1}{z^2}\sum_{r=0}^dR_r(z^{-1})(-\alpha_j+\partial_z)^r
\end{equation}

\color{brown}
Then taking the (formal) Borel transform:for every $j=1,...,d$ we denote $\zeta_j=\zeta-\alpha_j$ and we consider the Borel transform of $\mathcal{P}\series{\Psi}_j$ in the coordinate $\zeta_j$

\begin{multline*}
\mathcal{B}\left(\mathcal{P}\series{\Psi}_j\right)=\mathcal{B}\left(\mathcal{P}_j\sum_{n=0}^{\infty}a_{j,n}z^{-n-\tau_j}\right)=\\
\left[P(-\alpha_j-\zeta_j)+\partial_{\zeta_j}^{-1}\circ Q(-\alpha_j-\zeta_j)+\partial_{\zeta_j}^{-2}\circ \sum_{r=0}^dR_r(\partial_{\zeta_j}^{-1})\circ (-\alpha_j-\zeta_j)^r\right]\sum_{n=0}^{\infty}\frac{a_{j,n}}{\Gamma(n+\tau_j)}\zeta_j^{n+\tau_j-1}\\
\left[P(-\zeta)+\partial_{\zeta_j}^{-1}\circ Q(-\zeta)+\partial_{\zeta_j}^{-2}\circ \sum_{r=0}^dR_r(\partial_{\zeta_j}^{-1})\circ (-\zeta)^r\right]\sum_{n=0}^{\infty}\frac{a_{j,n}}{\Gamma(n+\tau_j)}\zeta_j^{n+\tau_j-1}
\end{multline*}
%where 
%
%\begin{align*}
%\series{\psi}_j(\zeta_j):=\mathcal{B}\series{\Psi}_j(\zeta_j)&=\mathcal{B}\left(e^{-\alpha_j z}z^{-\tau_j}\sum_{n=0}^{\infty}a_{j,n}z^{-n}\right)(\zeta_j)=\mathcal{B}\left(z^{-\tau_j}\sum_{n=0}^{\infty}a_{j,n}z^{-n}\right)(\zeta-\alpha_j)\\
%&=\sum_{n=0}^{\infty}\frac{a_{j,n}}{\Gamma(n+\tau_j)}(\zeta-\alpha_j)^{n+\tau_j-1}
%\end{align*}

Hence, the Borel transform of $\series{\Psi}_j$ is a solution of $\hat{P}_j\series{\psi}_j$=0.  from the properties of the Borel transform we deduce that $\hat{P}\series{\psi}_j=0$ first notice that $\series{\psi}_j$ is a solution of 
\color{black}
\end{proof}
